Supplementary Information for RVD2: An ultra-sensitive variant detection model for low-depth heterogeneous next-generation sequencing data

### 2015-03-09 doc/bioinformatics2014/revision2/response_to_reviews.txt

5. The authors greatly expose the comparison of sensibility, specificity, true positive and true negative rates. It seems that the caveat of the method is on memory consumption and
processing time. It would be very informative to explicitly have a table with execution time
and memory consumption of all utilized methods, preferably in a more realistic scenario,
more specifically, a whole genome analysis.

The memory consumption and computational time for the algorithm is an important consideration. We intentionally focused on the accuracy of the methodology here and so chose Metropolis-Hastings sampling as our inference algorithm. Metropolis-Hastings is known to converge to sampling from the true posterior distribution in the limit. As such, presently the algorithm is not as computationally fast as it could be. To improve the speed of the algorithm, we have implemented a multi-threaded version to take advantage of a multi-core computational configuration. We provide computational speed results in Section X.X and discuss avenues for improvement that we are currently pursuing.

Reviewer 2:

1. page 4 line 22; check the signs of the threshold boundary conditions in your hypothesis tests.. looks like one should be less than minus Tau..

Thank you for this observant correction. Indeed, the equation on page 4 line 22 should be minus tau. We have corrected the equation in the main text as well as the same equation in the supplementary information in the caption for Figure S6.

### 2015-03-09 doc/bioinformatics2014/revision2/document.tex
Responded to reviewer 2 comment 1 regarding -tau.

We use a two-sided posterior difference test with control and case paired samples to identify somatic mutations. We consider scenarios when the case(tumor) error rate is lower than the control(germline) error rate (e.g. loss-of-heterozygosity) as well as scenarios when the case(tumor) error rate is higher than the control(germline) error rate (e.g. homozygous somatic mutation). The two hypothesis tests are then $\Pr( \mu_j^{\text{case}} - \mu_j^{\text{control}} > \tau ) > 1-\alpha$ and $\Pr( \mu_j^{\text{case}} - \mu_j^{\text{control}} < \tau ) > 1-\alpha$. We typically set the threshold $ \tau $ to zero.

>>>>

We use a two-sided posterior difference test with control and case paired samples to identify somatic mutations. We consider scenarios when the case(tumor) error rate is lower than the control(germline) error rate (e.g. loss-of-heterozygosity) as well as scenarios when the case(tumor) error rate is higher than the control(germline) error rate (e.g. homozygous somatic mutation). The two hypothesis tests are then $\Pr( \mu_j^{\text{case}} - \mu_j^{\text{control}} > \tau ) > 1-\alpha$ and $\Pr( \mu_j^{\text{case}} - \mu_j^{\text{control}} < -\tau ) > 1-\alpha$. We typically set the threshold $ \tau $ to zero.

### 2015-03-09 doc/bioinformatics2014/revision2/supplementary_info.tex
Added pdf figures in docs folder and removed *.pdf from .gitignore.
Responded to reviewer 2 comment 1 regarding -tau.

The right column of Figure~\ref{fig:depthM} shows the parameter estimates $\hat{M}_j$ and $\hat{M}_0$ for each data set. $M_j$ measures the variance between replicates at location $j$. There is little variability across positions indicating that the replication variance does not change greatly across position. Furthermore, we see that $M_j$ does not change with read depth (except where the depth goes to zero) indicating that $M_j$ because $M_j$ is capturing a different process than the read depth.

@@

\begin{figure}[H]
\begin{center}
\includegraphics[width=1\textwidth]{pdf_figs/HCC1187_histogram_neg.pdf}
\caption{Histogram of $ \hat{\mu}_j $ for positions where $\mu^{\text{case}}$ is significantly lower than $ \mu^{\text{control}}$, namely Pr$ (\Delta \hat{\mu}_j< \tau) > 1- \alpha $, where $ \tau=0, \alpha=0.05 $.}
\label{fig:hist_neg}
\end{center}
\end{figure}

>>>>

The right column of Figure~\ref{fig:depthM} shows the parameter estimates $\hat{M}_j$ and $\hat{M}_0$ for each data set. $M_j$ measures the variance between replicates at location $j$. There is little variability across positions indicating that the replication variance does not change greatly across position.

%Furthermore, we see that $M_j$ does not change with read depth (except where the depth goes to zero) indicating that $M_j$ because $M_j$ is capturing a different process than the read depth.

@@

\begin{figure}[H]
\begin{center}
\includegraphics[width=1\textwidth]{pdf_figs/HCC1187_histogram_neg.pdf}
\caption{Histogram of $ \hat{\mu}_j $ for positions where $\mu^{\text{case}}$ is significantly lower than $ \mu^{\text{control}}$, namely Pr$ (\Delta \hat{\mu}_j< -\tau) > 1- \alpha $, where $ \tau=0, \alpha=0.05 $.}
\label{fig:hist_neg}
\end{center}
\end{figure}
